---
title: "hw5_fissel"
author: "Morgan Fissel"
date: "2/18/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Question 1. 
a) Perform a likelihood ratio test at significance level 0.04 on the melanoma$thickness data.
The assumption is that the data are Gamma distributed and the null hypothesis is that the
shape parameter α is 1 and the alternative hypothesis is that it is not.
```{r}
# data - the package 'boot' needs to be installed
library(boot)
data(melanoma); y = melanoma$thickness
#density - dgamma - built in
# mle
mle_gamma = function(y,vcov=TRUE){
  #takes advantage of scoping
  loglik_fun = function(theta){#scoping, looks for y one environment up
    alpha = theta[1]
    beta = theta[2]
    sum(dgamma(y,shape=alpha,rate=beta,log=TRUE))
  }
  loglik_fun_trans = function(theta){#scoping, looks for y one environment up
    alpha = exp(theta[1])
    beta = exp(theta[2])
    sum(dgamma(y,shape=alpha,rate=beta,log=TRUE))
  }
  #using optim to get MLE
  #we use the _trans function and then undo the transformation for the estimate without the transformation
  fit_trans = optim(c(0,0),loglik_fun_trans,control=list(fnscale=-1,maxit=10000)) #scoping, looks for y one environment up
  fit = fit_trans
  fit$par[1] = exp(fit$par[1])
  fit$par[2] = exp(fit$par[2])
  dim = 2
  AIC = -2*fit$value + 2*dim
  names(fit$par) = c("alpha","beta") #NAMES ARE GOOD
  if(!vcov){
    return(list(theta=fit$par,loglik=fit$value,dim=dim,AIC=AIC))
  }else{
    #using optimHess to get the Hessian
    #we use the regular log-likelihood function
    fit_hess = optimHess(fit$par,loglik_fun)#scoping, looks for y one environment up
    #we convert the Hessian into a variance-covariance matrix
    #inversion is achieved using solve
    fit_vcov = solve(-fit_hess)
    rownames(fit_vcov) = colnames(fit_vcov) = names(fit$par)#NAMES ARE COPIED HERE
    return(list(theta=fit$par,loglik=fit$value,dim=dim,AIC=AIC,vcov=fit_vcov))
  }
}
```

```{r}
# now null
mle_gamma_null = function(y,vcov=TRUE){
  #takes advantage of scoping
  loglik_fun = function(theta){#scoping, looks for y one environment up
    alpha = 1 #theta[1]
    beta = theta[1]
    sum(dgamma(y,shape=alpha,rate=beta,log=TRUE))
  }
  loglik_fun_trans = function(theta){#scoping, looks for y one environment up
    alpha = 1 #exp(theta[1])
    beta = exp(theta[1])
    sum(dgamma(y,shape=alpha,rate=beta,log=TRUE))
  }
  #using optim to get MLE
  #we use the _trans function and then undo the transformation for the estimate without the transformation
  ##AJW## one dimensional optimization instead
  fit_trans = optim(c(0),loglik_fun_trans,
                    method="Brent", lower=-100, upper=100,
                    control=list(fnscale=-1,maxit=10000)) #scoping, looks for y one environment up
  fit = fit_trans
  fit$par[1] = exp(fit$par[1])
  ##AJW## one dimensional parameter
  #fit$par[2] = exp(fit$par[2])
  dim = 1
  AIC = -2*fit$value + 2*dim
  ##AJW## one dimensional parameter
  names(fit$par) = "beta" #c("alpha","beta") #NAMES ARE GOOD
  
  return(list(theta=fit$par,loglik=fit$value,dim=dim,AIC=AIC))
  
  ##AJW## returned already, none of the below is seen
  #using optimHess to get the Hessian
  #we use the regular log-likelihood function
  fit_hess = optimHess(fit$par,loglik_fun)#scoping, looks for y one environment up
  #we convert the Hessian into a variance-covariance matrix
  #inversion is achieved using solve
  fit_vcov = solve(-fit_hess)
  rownames(fit_vcov) = colnames(fit_vcov) = names(fit$par)#NAMES ARE COPIED HERE
  return(list(theta=fit$par,loglik=fit$value,dim=dim,AIC=AIC,vcov=fit_vcov))
  
}
```

```{r}
# make a significance level for the test
significance_level = 0.04
  
# LRT computations
fit_alternative = mle_gamma(y)
fit_null = mle_gamma_null(y)

test_stat = 2*(fit_alternative$loglik - fit_null$loglik)
print(test_stat)
df = fit_alternative$dim - fit_null$dim
p_value = pchisq(test_stat,df=df ,lower.tail = FALSE)
print(p_value)
cutoff = qchisq(significance_level,df=df,lower.tail=FALSE)
print(cutoff)
if(test_stat>cutoff){
  decision = 'Reject H0'
}else{
  decision = 'Fail to reject H0'
}
print(decision)
```
b) Perform a likelihood ratio test at significance level 0.03 on the log of the melanoma$thickness
data. The assumption is that the data are Logistic Type III distributed and the null hypothesis
is that the shape parameter α is 1 and the alternative hypothesis is that it is not.
```{r}
# data - the package 'boot' needs to be installed
library(boot); data(melanoma); y = log(melanoma$thickness)
#density - dlogisIII - provided
dlogisIII=function(x,location=0,scale=1,shape=1,log=FALSE){
  x = (x-location)/scale
  log_y = ifelse(x<0,x-log(1+exp(x)),-log(1+exp(-x)))
  log_1my = ifelse(x<0,-log(1+exp(x)),-x-log(1+exp(-x)))
  out = -log(scale)-lbeta(shape,shape) + (shape)*(log_y+log_1my)
  if(log){return(out)}else{return(exp(out))}
}
# mle
mle_logisIII = function(y,vcov=TRUE){
  #takes advantage of scoping
  loglik_fun = function(theta){#scoping, looks for y one environment up
    mu = theta[1]
    delta = theta[2]
    alpha = theta[3]
    sum(dlogisIII(y,location=mu,scale=delta,shape=alpha,log=TRUE))
  }
  loglik_fun_trans = function(theta){#scoping, looks for y one environment up
    mu = theta[1]
    delta = exp(theta[2])
    alpha = exp(theta[3])
    sum(dlogisIII(y,location=mu,scale=delta,shape=alpha,log=TRUE))
  }
  #using optim to get MLE
  #we use the _trans function and then undo the transformation for the estimate without the transformation
  fit_trans = optim(c(0,0,0),loglik_fun_trans,control=list(fnscale=-1,maxit=10000)) #scoping, looks for y one environment up
  fit = fit_trans
  fit$par[1] = fit$par[1]
  fit$par[2] = exp(fit$par[2])
  fit$par[3] = exp(fit$par[3])
  dim = 3
  AIC = -2*fit$value + 2*dim
  names(fit$par) = c("mu","delta","alpha") #NAMES ARE GOOD
  if(!vcov){
    return(list(theta=fit$par,loglik=fit$value,dim=dim,AIC=AIC))
  }else{
    #using optimHess to get the Hessian
    #we use the regular log-likelihood function
    fit_hess = optimHess(fit$par,loglik_fun)#scoping, looks for y one environment up
    #we convert the Hessian into a variance-covariance matrix
    #inversion is achieved using solve
    fit_vcov = solve(-fit_hess)
    rownames(fit_vcov) = colnames(fit_vcov) = names(fit$par)#NAMES ARE COPIED HERE
    return(list(theta=fit$par,loglik=fit$value,dim=dim,AIC=AIC,vcov=fit_vcov))
  }
}
```

```{r}
mle_logisIII_null = function(y,vcov=TRUE){
  #takes advantage of scoping
  loglik_fun = function(theta){#scoping, looks for y one environment up
    mu = theta[1]
    delta = theta[2]
    alpha = 1 #theta[3]
    sum(dlogisIII(y,location=mu,scale=delta,shape=alpha,log=TRUE))
  }
  loglik_fun_trans = function(theta){#scoping, looks for y one environment up
    mu = theta[1]
    delta = exp(theta[2])
    alpha = 1 #exp(theta[3])
    sum(dlogisIII(y,location=mu,scale=delta,shape=alpha,log=TRUE))
  }
  #using optim to get MLE
  #we use the _trans function and then undo the transformation for the estimate without the transformation
  fit_trans = optim(c(0,0),loglik_fun_trans,control=list(fnscale=-1,maxit=10000)) #scoping, looks for y one environment up
  fit = fit_trans
  ##AJW## two dimensional parameter
  fit$par[1] = fit$par[1]
  fit$par[2] = exp(fit$par[2])
  #fit$par[3] = exp(fit$par[3])
  dim = 2
  AIC = -2*fit$value + 2*dim
  ##AJW## two dimensional parameter
  names(fit$par) = c("mu","delta")#,"alpha") #NAMES ARE GOOD
  
  return(list(theta=fit$par,loglik=fit$value,dim=dim,AIC=AIC))
  
  ##AJW## returned already, none of the below is seen
  #using optimHess to get the Hessian
  #we use the regular log-likelihood function
  fit_hess = optimHess(fit$par,loglik_fun)#scoping, looks for y one environment up
  #we convert the Hessian into a variance-covariance matrix
  #inversion is achieved using solve
  fit_vcov = solve(-fit_hess)
  rownames(fit_vcov) = colnames(fit_vcov) = names(fit$par)#NAMES ARE COPIED HERE
  return(list(theta=fit$par,loglik=fit$value,dim=dim,AIC=AIC,vcov=fit_vcov))
  
}
```

```{r}
# make a significance level for the test
significance_level = 0.03
  
# LRT computations
fit_alternative = mle_logisIII(y)
fit_null = mle_logisIII_null(y)

test_stat = 2*(fit_alternative$loglik - fit_null$loglik)
print(test_stat)
df = fit_alternative$dim - fit_null$dim
p_value = pchisq(test_stat,df=df ,lower.tail = FALSE)
print(p_value)
cutoff = qchisq(significance_level,df=df,lower.tail=FALSE)
print(cutoff)
if(test_stat>cutoff){
  decision = 'Reject H0'
}else{
  decision = 'Fail to reject H0'
}
print(decision)
```
c) Perform a likelihood ratio test at significance level 0.05 on the a subset of the faithful$eruptions
data subsetted for the observations where faithful$waiting>71. The assumption is that the
data are Generalized-Gamma distributed and the null hypothesis is the Weibull distribution.
The null hypothesis is that the shape parameters d and k are equal and the alternative hypothesis is that they are not.
```{r}
y = faithful$eruptions[faithful$waiting>71]
#density - dgengamma - provided
dgengamma = function(y,d,k,s,log=FALSE){
  out = log(k) -lgamma(d/k) -d*log(s) +log(y)*(d-1) -(y/s)^k
  if(log){out}else{exp(out)}
}
# mle
mle_gengamma = function(y,vcov=TRUE){
  #takes advantage of scoping
  loglik_fun = function(theta){#scoping, looks for y one environment up
    d = theta[1]
    k = theta[2]
    s = theta[3]
    sum(dgengamma(y,d,k,s,log=TRUE))
  }
  loglik_fun_trans = function(theta){#scoping, looks for y one environment up
    d = exp(theta[1])
    k = exp(theta[2])
    s = exp(theta[3])
    sum(dgengamma(y,d,k,s,log=TRUE))
  }
  #using optim to get MLE
  #we use the _trans function and then undo the transformation for the estimate without the transformation
  fit_trans = optim(c(0,0,0),loglik_fun_trans,control=list(fnscale=-1,maxit=10000)) #scoping, looks for y one environment up
  fit = fit_trans
  fit$par[1] = exp(fit$par[1])
  fit$par[2] = exp(fit$par[2])
  fit$par[3] = exp(fit$par[3])
  dim = 3
  AIC = -2*fit$value + 2*dim
  names(fit$par) = c("d","k","s") #NAMES ARE GOOD
  return(list(theta=fit$par,loglik=fit$value,dim=dim,AIC=AIC))
  
  #using optimHess to get the Hessian
  #we use the regular log-likelihood function
  fit_hess = optimHess(fit$par,loglik_fun)#scoping, looks for y one environment up
  #we convert the Hessian into a variance-covariance matrix
  #inversion is achieved using solve
  fit_vcov = solve(-fit_hess)
  rownames(fit_vcov) = colnames(fit_vcov) = names(fit$par)#NAMES ARE COPIED HERE
  return(list(theta=fit$par,loglik=fit$value,dim=dim,AIC=AIC,vcov=fit_vcov))
}

```

```{r}
mle_weibull_null = function(y,vcov=TRUE){
  #takes advantage of scoping
  loglik_fun = function(theta){#scoping, looks for y one environment up
    ##AJW## still need scale
    ## you are using dweibull instead of restricting dgengamma
    ##scale is s, and dweibull has no second shape parameter
    d = theta[1]
    k = d
    s=theta[2]#lambda = theta[2]
    sum(dweibull(y,shape=k,scale=s,log=TRUE))
  }
  loglik_fun_trans = function(theta){#scoping, looks for y one environment up
    ##AJW## still need scale
    ## you are using dweibull instead of restricting dgengamma
    ##scale is s, and dweibull has no second shape parameter
    k = exp(theta[1])
    d = k
    s=exp(theta[2])#lambda = exp(theta[2])
    sum(dweibull(y,shape=k,scale=s,log=TRUE))
  }
  #using optim to get MLE
  #we use the _trans function and then undo the transformation for the estimate without the transformation
  fit_trans = optim(c(0,0),loglik_fun_trans,control=list(fnscale=-1)) #scoping, looks for y one environment up
  fit = fit_trans
  fit$par[1] = exp(fit$par[1])
  ##AJW##
  ##get scale out
  #fit$par[2] = fit$par[1]#exp(fit$par[2])
  fit$par[2] = exp(fit$par[2])
  dim = 2
  AIC = -2*fit$value + 2*dim
  ##AJW## scale s is second parameter
  names(fit$par) = c("k","s") #NAMES ARE GOOD
  return(list(theta=fit$par,loglik=fit$value,dim=dim,AIC=AIC))
  
  #using optimHess to get the Hessian
  #we use the regular log-likelihood function
  fit_hess = optimHess(fit$par,loglik_fun)#scoping, looks for y one environment up
  #we convert the Hessian into a variance-covariance matrix
  #inversion is achieved using solve
  fit_vcov = solve(-fit_hess)
  rownames(fit_vcov) = colnames(fit_vcov) = names(fit$par)#NAMES ARE COPIED HERE
  return(list(theta=fit$par,loglik=fit$value,vcov=fit_vcov))
}

```

```{r}
# make a significance level for the test
significance_level = 0.05
  
# LRT computations
fit_alternative = mle_gengamma(y)
fit_null = mle_weibull_null(y)

test_stat = 2*(fit_alternative$loglik - fit_null$loglik)
print(test_stat)
df = fit_alternative$dim - fit_null$dim
p_value = pchisq(test_stat,df=df ,lower.tail = FALSE)
print(p_value)
cutoff = qchisq(significance_level,df=df,lower.tail=FALSE)
print(cutoff)
if(test_stat>cutoff){
  decision = 'Reject H0'
}else{
  decision = 'Fail to reject H0'
}
print(decision)
```
d) Perform a likelihood ratio test at significance level 0.01 on the acme$acme data. The assumption is that the data are Hyperbolically distributed and the null hypothesis is that the mean
parameter µ is 0 and the alternative hypothesis is that it is not.
```{r}
# data - the package 'boot' needs to be installed
library(boot); data("acme"); y = acme$acme
#density - dhyperbolic - provided
dhyperbolic = function(y,mu=0,delta=1,alpha=1,log=FALSE){
  if(log(alpha)< -20){
    log_NC = -log(alpha)+log(2)+log(delta)
  }else{
    log_NC = log(besselK(alpha,1,TRUE)) - alpha + log(2)+log(delta)
  }
  loss = alpha*sqrt(1+(y-mu)^2/delta^2)
  out = -loss -log_NC
  if(log==TRUE){out}else{exp(out)}
}
# mle
mle_hyperbolic = function(y,vcov=TRUE){
  #takes advantage of scoping
  loglik_fun = function(theta){#scoping, looks for y one environment up
    mu = theta[1]
    delta = theta[2]
    alpha = theta[3]
    sum(dhyperbolic(y,mu,delta,alpha,log=TRUE))
  }
  loglik_fun_trans = function(theta){#scoping, looks for y one environment up
    mu = theta[1]
    delta = exp(theta[2])
    alpha = exp(theta[3])
    sum(dhyperbolic(y,mu,delta,alpha,log=TRUE))
  }
  #using optim to get MLE
  #we use the _trans function and then undo the transformation for the estimate without the transformation
  fit_trans = optim(c(0,0,0),loglik_fun_trans,control=list(fnscale=-1,maxit=10000)) #scoping, looks for y one environment up
  fit = fit_trans
  fit$par[1] = fit$par[1]
  fit$par[2] = exp(fit$par[2])
  fit$par[3] = exp(fit$par[3])
  dim = 3
  AIC = -2*fit$value + 2*dim
  names(fit$par) = c("mu","delta","alpha") #NAMES ARE GOOD
  if(!vcov){
    return(list(theta=fit$par,loglik=fit$value,dim=dim,AIC=AIC))
  }else{
    #using optimHess to get the Hessian
    #we use the regular log-likelihood function
    fit_hess = optimHess(fit$par,loglik_fun)#scoping, looks for y one environment up
    #we convert the Hessian into a variance-covariance matrix
    #inversion is achieved using solve
    fit_vcov = solve(-fit_hess)
    rownames(fit_vcov) = colnames(fit_vcov) = names(fit$par)#NAMES ARE COPIED HERE
    return(list(theta=fit$par,loglik=fit$value,dim=dim,AIC=AIC,vcov=fit_vcov))
  }
}
```

```{r}
mle_hyperbolic_null = function(y,vcov=TRUE){
  #takes advantage of scoping
  loglik_fun = function(theta){#scoping, looks for y one environment up
    mu = 0#theta[1]
    delta = theta[1]
    alpha = theta[2]
    sum(dhyperbolic(y,mu,delta,alpha,log=TRUE))
  }
  loglik_fun_trans = function(theta){#scoping, looks for y one environment up
    mu = 0#theta[1]
    delta = exp(theta[1])
    alpha = exp(theta[2])
    sum(dhyperbolic(y,mu,delta,alpha,log=TRUE))
  }
  #using optim to get MLE
  #we use the _trans function and then undo the transformation for the estimate without the transformation
  fit_trans = optim(c(0,0),loglik_fun_trans,control=list(fnscale=-1,maxit=10000)) #scoping, looks for y one environment up
  fit = fit_trans
  fit$par[1] = fit$par[1]
  fit$par[2] = exp(fit$par[2])
  fit$par[3] = exp(fit$par[3])
  dim = 2
  AIC = -2*fit$value + 2*dim
  names(fit$par) = c("mu","delta","alpha") #NAMES ARE GOOD
  
  return(list(theta=fit$par,loglik=fit$value,dim=dim,AIC=AIC))
  
  #using optimHess to get the Hessian
  #we use the regular log-likelihood function
  fit_hess = optimHess(fit$par,loglik_fun)#scoping, looks for y one environment up
  #we convert the Hessian into a variance-covariance matrix
  #inversion is achieved using solve
  fit_vcov = solve(-fit_hess)
  rownames(fit_vcov) = colnames(fit_vcov) = names(fit$par)#NAMES ARE COPIED HERE
  return(list(theta=fit$par,loglik=fit$value,dim=dim,AIC=AIC,vcov=fit_vcov))
 
}
```
```{r}
# make a significance level for the test
significance_level = 0.01
  
# LRT computations
fit_alternative = mle_hyperbolic(y)
fit_null = mle_hyperbolic_null(y)

test_stat = 2*(fit_alternative$loglik - fit_null$loglik)
print(test_stat)
df = fit_alternative$dim - fit_null$dim
p_value = pchisq(test_stat,df=df ,lower.tail = FALSE)
print(p_value)
cutoff = qchisq(significance_level,df=df,lower.tail=FALSE)
print(cutoff)
if(test_stat>cutoff){
  decision = 'Reject H0'
}else{
  decision = 'Fail to reject H0'
}
print(decision)
```

e) Perform a likelihood ratio test at significance level 0.001 on the penguins data where the response variable is y=penguins\$flipper_length_mm and the covariate is penguins$body_mass_g.
The assumption is that the data follow a Gaussian Linear Model and the null hypothesis is
that the slope parameter β is 0 and the alternative hypothesis is that it is not.
```{r}
# data - the package 'palmerpenguins' needs to be installed
library(palmerpenguins)
y = penguins$flipper_length_mm
x = penguins$body_mass_g
ind = !(is.na(x)&is.na(y))
y=y[ind]
x=x[ind]
#density - dlinmod - provided
dlinmod = function(x,y,alpha,beta,sigma,log=FALSE){
  dnorm(y,alpha+beta*x,sigma,log)
}
# mle
mle_linmod = function(x,y,vcov=TRUE){
  #takes advantage of scoping
  loglik_fun = function(theta){#scoping, looks for y one environment up
    alpha = theta[1]
    beta = theta[2]
    sigma = theta[3]
    sum(dlinmod(x,y,alpha,beta,sigma,log=TRUE))
  }
  loglik_fun_trans = function(theta){#scoping, looks for y one environment up
    alpha = theta[1]
    beta = theta[2]
    sigma = exp(theta[3])
    sum(dlinmod(x,y,alpha,beta,sigma,log=TRUE))
  }
  #using optim to get MLE
  #we use the _trans function and then undo the transformation for the estimate without the transformation
  #I had to change the starting point for the optimization for optim to work
  fit_trans = optim(c(mean(y),0,0),loglik_fun_trans,control=list(fnscale=-1,maxit=10000)) #scoping, looks for y one environment up
  fit = fit_trans
  fit$par[1] = fit$par[1]
  fit$par[2] = fit$par[2]
  fit$par[3] = exp(fit$par[3])
  dim = 3
  AIC = -2*fit$value + 2*dim
  names(fit$par) = c("alpha","beta","sigma") #NAMES ARE GOOD
  if(!vcov){
    return(list(theta=fit$par,loglik=fit$value,dim=dim,AIC=AIC))
  }else{
    #using optimHess to get the Hessian
    #we use the regular log-likelihood function
    fit_hess = optimHess(fit$par,loglik_fun)#scoping, looks for y one environment up
    #we convert the Hessian into a variance-covariance matrix
    #inversion is achieved using solve
    fit_vcov = solve(-fit_hess)
    rownames(fit_vcov) = colnames(fit_vcov) = names(fit$par)#NAMES ARE COPIED HERE
    return(list(theta=fit$par,loglik=fit$value,dim=dim,AIC=AIC,vcov=fit_vcov))
  }
}
```

Now null
```{r}
mle_linmod_null = function(x,y,vcov=TRUE){
  #takes advantage of scoping
  loglik_fun = function(theta){#scoping, looks for y one environment up
    alpha = theta[1]
    beta = 0#theta[2]
    sigma = theta[2]
    sum(dlinmod(x,y,alpha,beta,sigma,log=TRUE))
  }
  loglik_fun_trans = function(theta){#scoping, looks for y one environment up
    alpha = theta[1]
    beta = 0#theta[2]
    sigma = exp(theta[2])
    sum(dlinmod(x,y,alpha,beta,sigma,log=TRUE))
  }
  #using optim to get MLE
  #we use the _trans function and then undo the transformation for the estimate without the transformation
  #I had to change the starting point for the optimization for optim to work
  fit_trans = optim(c(mean(y),0),loglik_fun_trans,control=list(fnscale=-1,maxit=10000)) #scoping, looks for y one environment up
  fit = fit_trans
  ##AJW##
  # two dimensional parameter
  fit$par[1] = fit$par[1]
  fit$par[2] = exp(fit$par[2])
  dim = 2
  AIC = -2*fit$value + 2*dim
  ##AJW##
  # two dimensional parameter
  names(fit$par) = c("alpha","sigma") #c("alpha","beta","sigma") #NAMES ARE GOOD
  
  return(list(theta=fit$par,loglik=fit$value,dim=dim,AIC=AIC))
  
  #using optimHess to get the Hessian
  #we use the regular log-likelihood function
  fit_hess = optimHess(fit$par,loglik_fun)#scoping, looks for y one environment up
  #we convert the Hessian into a variance-covariance matrix
  #inversion is achieved using solve
  fit_vcov = solve(-fit_hess)
  rownames(fit_vcov) = colnames(fit_vcov) = names(fit$par)#NAMES ARE COPIED HERE
  return(list(theta=fit$par,loglik=fit$value,dim=dim,AIC=AIC,vcov=fit_vcov))
  
}
```
```{r}
# make a significance level for the test
significance_level = 0.001
  
# LRT computations
fit_alternative = mle_linmod(x,y)
fit_null = mle_linmod_null(x,y)

test_stat = 2*(fit_alternative$loglik - fit_null$loglik)
print(test_stat)
df = fit_alternative$dim - fit_null$dim
p_value = pchisq(test_stat,df=df ,lower.tail = FALSE)
print(p_value)
cutoff = qchisq(significance_level,df=df,lower.tail=FALSE)
print(cutoff)
if(test_stat>cutoff){
  decision = 'Reject H0'
}else{
  decision = 'Fail to reject H0'
}
print(decision)
```
