---
title: "hw3_fissel"
author: "Morgan Fissel"
date: "2/9/2022"
output:
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
# load melanoma tumor thickness dataset
library(boot); data(melanoma); y = log(melanoma$thickness)
```

Helper Functions
```{r}
##AJW## just moved this up so it is in memory before we use dlogisIII
dlogisIII=function(x,location=0,scale=1,shape=1,log=FALSE){
  x = (x-location)/scale
  log_y = ifelse(x<0,x-log(1+exp(x)),-log(1+exp(-x)))
  log_1my = ifelse(x<0,-log(1+exp(x)),-x-log(1+exp(-x)))
  out = -log(scale)-lbeta(shape,shape) + (shape)*(log_y+log_1my)
  if(log){return(out)}else{return(exp(out))}
}
plogisIII=function(q,location=0,scale=1,shape=1,lower.tail=TRUE,log.p=FALSE){
  q = (q-location)/scale
  Q = ifelse(q<0,exp(q)/(1+exp(q)),1/(1+exp(-q)))
  return(pbeta(Q,shape,shape,0,lower.tail,log.p))
}
qlogisIII=function(p,location=0,scale=1,shape=1,lower.tail=TRUE,log.p=FALSE){
  Q = qbeta(p,shape,shape,0,lower.tail,log.p)
  q = ifelse(Q<0.5,log(Q)-log(1-Q),-log(1/Q-1))
  return(q*scale+location)
}
rlogisIII=function(n,location=0,scale=1,shape=1){
  b = rbeta(n,shape,shape,0)
  x = ifelse(b<0.5,log(b)-log(1-b),-log(1/b-1))
  return(x*scale+location)
}

```
a) Write MLE Function like from likelihoods.pdf. 
Will use to Compute MLE for Type III Logisitic model. 
Note: scaling delta and shape alpha are positive.
```{r}
mle_logisIII = function(y){
  #function to optimize using optim
  f = function(theta){
    mu = theta[1]
    delta = exp(theta[2]) #note the transformation
    ##AJW## there are three parameters
    alpha = exp(theta[3])
    #sum(dlogisIII(y,location=mu,scale=delta,log=TRUE))
    sum(dlogisIII(y,location=mu,scale=delta,shape=alpha,log=TRUE))
  }
  #using optim - starting at a guess of mu=0 and log(delta)=0 -- par=c(0,0)
  ##AJW## 3 dimensional optimization, need 3 dimensional starting point
  #opt_out = optim(par=c(0,0),f,control=list(fnscale=-1,maxit=1000,reltol=1e-8))
  opt_out = optim(par=c(0,0,0),f,control=list(fnscale=-1,maxit=1000,reltol=1e-8))
  par = opt_out$par
  par[2] = exp(par[2]) #transforming to delta
  ##AJW## geting alpha out and adding name to names
  par[3] = exp(par[3])
  #names(par) = c("mu","delta") #NAMES ARE GOOD
  names(par) = c("mu","delta","alpha") #NAMES ARE GOOD
  out = list(theta=par,loglik=opt_out$value)
  return(out)
}

##AJW## need to make a funciton for logistic type III,
## it should use the helper functions below and be a modification of this function
## to do the 3 parameter optimization
```

b) Compute the mle_logis(y=log(melanoma$thickness))
Plot density of the model using dlogisIII with KDE of y
Does the model appear to capture the data well?
```{r}
mle_logisIII(y)
fit_logisIII_y = mle_logisIII(y)

plot(density(y),xlim=c(-3,4),ylim=c(0,0.45))

##AJW## some changes here
##I am going to comment your line and do what the line would be for the for
##to the logistic distribution (shape parameter alpha=1)
#lines(dlogisIII(y,fit_logis_y[["theta"]][["mu"]],fit_logis_y[["theta"]][["delta"]],1))
x_val = seq(-3,4,length.out=1000)
##AJW## there are 3 parameters
#dens_logisIII_vals = dlogisIII(x_val,fit_logisIII_y[["theta"]][["mu"]],fit_logisIII_y[["theta"]][["delta"]])
dens_logisIII_vals = dlogisIII(x_val,fit_logisIII_y[["theta"]]["mu"],fit_logisIII_y[["theta"]]["delta"],fit_logisIII_y[["theta"]]["alpha"])
lines(dens_logisIII_vals~x_val,col="blue",lty=2)
legend("bottom",legend=c("KDE","Fitted"),col=c("black","blue"),lty=c(1,2),bty="n")


```
The Model does a decent job at capturing the data. 
The middle of the fitted graph is a bit taller than the KDE, however it smoothes the curves of the KDE quite nicely. 

c) Do a goodness of fit test for the fitted Type III logisitic distribution using ks.test. 
```{r}
##AJW##
##because your fit from above is the the logistic and not logistic type 3, 
##I am going to rewrite this line to demonstrate comparison to logistic
##your setting alpha=1 is doing that, but I think it will be cleaner to see this again
##ks.test(y, dlogisIII, location = fit_logis_y[["theta"]][["mu"]], 
##        scale = fit_logis_y[["theta"]][["delta"]], shape = 1)
## comparison is to a probability function and not a density function
## so plogis instead of dlogis
##AJW## logistic III, so need all 3 parameters
#ks.test(y, plogisIII, location = fit_logisIII_y[["theta"]][["mu"]], 
#        scale = fit_logisIII_y[["theta"]][["delta"]])
ks.test(y, plogisIII, location = fit_logisIII_y[["theta"]]["mu"], 
        scale = fit_logisIII_y[["theta"]]["delta"],
        shape = fit_logisIII_y[["theta"]]["alpha"])
```
d) Perform tasks b) and c) using the mle_logis function that I provided and the built-in dlogis
and plogis functions.
```{r}
mle_logis = function(y){
  #function to optimize using optim
  f = function(theta){
    mu = theta[1]
    delta = exp(theta[2]) #note the transformation
    sum(dlogis(y,location=mu,scale=delta,log=TRUE))
  }
  #using optim - starting at a guess of mu=0 and log(delta)=0 -- par=c(0,0)
  opt_out = optim(par=c(0,0),f,control=list(fnscale=-1,maxit=1000,reltol=1e-8))
  par = opt_out$par
  par[2] = exp(par[2]) #transforming to delta
  names(par) = c("mu","delta") #NAMES ARE GOOD
  out = list(theta=par,loglik=opt_out$value)
  return(out)
}
mle_logis(y)
fit_logis_y = mle_logis(y)

dens_y = density(y)
x_seq = dens_y$x #seq(lb, ub,length.out=whatever)
fitted_dens_y = dlogis(x_seq,fit_logis_y$theta['mu'],fit_logis_y$theta['delta'])

plot(density(y),main='model fit plot',xlab='y',ylab='desity',ylim=c(0,max(c(dens_y$y,fitted_dens_y))),xlim=range(x_seq))
lines(fitted_dens_y~x_seq,lty=1,col='purple')
legend("bottom",legend=c("KDE","Fitted"),col=c("black","purple"),lty=(1),bty="n")

###

ks.test(y, plogis, location = fit_logis_y[["theta"]][["mu"]], 
        scale = fit_logis_y[["theta"]][["delta"]])

```
I'd say the model does a decent job handling the data.
The center is taller, and has much smoother curve.

e) Compare the log-likelihood evaluated at the MLE for the Logistic and Type III logistic models.
Which model has higher log-likelihood? Which model is better at capturing the data? Do you
think it was worth adding the shape parameter α to the Logistic distribution for modeling
this dataset? 
The Logisitic had a slightly higher loglikelihood at the MLE than the Type III.
I would say the Type III model is better at capturing the data. It more accurately mimics the shape of the KDE.
This is likely because of the alpha parameter, allowing us to better capture the dataset. 
Absolutely worth adding the alpha, I think.