---
title: "hw10_fissel"
author: "Morgan Fissel 2000498470"
date: "4/8/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Question 2
a)
```{r}
library(boot)
data(nodal)
y = nodal$stage
n = length(y)
alpha0 = 1
beta0 = 1
sn = sum(y) # successes
sn
fn = n - sn # fails 
fn
alpha_n = sn + alpha0 
beta_n = fn + beta0

N_samps = 10000
draws = rbeta(N_samps,alpha_n,beta_n)

plot(density(draws))
mean(draws)
quantile(draws,c(0.025, 0.975))

y_new = rbinom(N_samps,1,draws)
```
The model's predictions look good to me. Given there are 27 successes and 26 failures, and the probability should be about .50, we can conclude that having nearly equal successes and failures likely means we have about 50% probability of success.

b)
```{r}
library(boot)
data(fir)
y = fir$count
n = length(y)
sn = sum(y)
mean(y)
alpha0 = 1 
beta0 = 1
alpha_n = sn + alpha0
beta_n = n + beta0

N_samps = 10000 
draws = rgamma(N_samps,alpha_n,beta_n)

plot(density(draws))
mean(draws)
quantile(draws,c(0.025,0.975))

y_new = rpois(N_samps, draws)
```
Once again, our model is doing a decent job, the mean is about 2.118352, and our data's is about 2.14. Most of our numbers are going to be from like 1.7 to about 2.5. Which is exactly what our model shows.

c)
```{r}
library(boot)
data(poisons)
y = poisons$time
shape = 4.31 #use fixed shape
sn = sum(y)
mean(y)
alpha0 = 1
beta0 = 1
n = length(y)
alpha_n = n*shape + alpha0
beta_n = sn + beta0
N_samps = 10000

draws = rgamma(N_samps,alpha_n,beta_n)
plot(density(draws))
mean(draws)

quantile(draws,c(0.025,0.975))
y_new = rgamma(N_samps,shape,draws)
plot(density(y_new))
```
This model does not seem to represent our data well, it has a different mean entirely, and the shapes of the densities aren't very similar.

d)
```{r}
library(boot)
data(acme)
y=acme$acme
s_n = sum(y)
un = sum(y^2)
n = length(y)
N_samps = 10000
a_curr=0
b_curr=1
delta0=1
zeta0=1
mu_curr=0

## simple Inverting of draws from appropriate gamma distribution
## parameterized using shape = delta and scale = zeta
rinvgamma = function(n,delta,zeta){
  shape = delta0
  rate = zeta0
  1/rgamma(n,shape,rate)
}

#making containers for output
mu_out = a_out = b_out = rep(NA,N_samps)

#Looping for the sampler
for(i in 1:N_samps){
  #updating
  b_curr = rinvgamma(1,0.5+10,0.5*(mu_curr-a_curr)^2+1)
  a_curr = rnorm(1,(mu_curr/b_curr+0)/(1/b_curr+1),sqrt(1/(1/b_curr+1)))
  mu_curr = rnorm(1,(s_n+a_curr/b_curr)/(n+1/b_curr),sqrt(1/(n+1/b_curr)))
  #storing
  mu_out[i] = mu_curr
  a_out[i] = a_curr
  b_out[i] = b_curr
}

#looking at the output
plot(mu_out,pch='.')
acf(mu_out)
plot(a_out,pch='.')
acf(a_out)
plot(b_out,pch='.')
acf(b_out)

#Drawing inference
#Density plots
plot(density(mu_out))
plot(density(a_out))
plot(density(b_out,from=0))
#posterior means
mean(mu_out)
mean(a_out)
mean(b_out)
#intervals
quantile(mu_out,c(0.025,0.975))
quantile(a_out,c(0.025,0.975))
quantile(b_out,c(0.025,0.975))
#correlation of parameters
cor(mu_out,a_out)
cor(mu_out,b_out)
cor(a_out,b_out)


#predictions
y_pred = rnorm(N_samps,mu_out,1)
#comparison
plot(density(y))
lines(density(y_pred),col="blue",lty=2)
legend("topright",legend=c("KDE-Data","KDE-Predictions"),col=c("black","blue"),lty=c(1,2))


##
## gibbs example with three random variables
## the model for the data is y_i|mu~N(mu,1)
## the prior for mu|a,b is mu~N(a,b)
## a and b are given a prior 
## and assumed independent in the prior
## the prior for a is a~N(0,1)
## the prior for b is b~InverseGamma(10,1)
## The full conditional for mu is N((s_n+a/b)/(n+1/b),1/(n+1/b))
## The full conditional for a is N((mu/b+0)/(1/b+1),1/(1/b+1))
## The full conditional for b is InverseGamma(0.5+10,0.5*(mu-a)^2+1)
```

The predictions don't match the data very well. The model is likely insufficient for the data. The trace and ACF plots are generally not great, but could possibly work. I may be doing something wrong here. a_out should be our nu_out, or v_out (shape).

e)
```{r, eval=FALSE}
library(palmerpenguins)
data("penguins")
y = penguins$flipper_length_mm
x = penguins$body_mass_g
ind = !(is.na(y)|is.na(x))
y = y[ind]
x = x[ind]

n = length(y)
s_n = sum(y)
u_n = sum(y^2)
t_n = sum(x)
r_n = sum(x*y)
w_n = sum(x^2)

a_0=0
b_0=1e6
c_0=0
d_0=1
zeta_0 = 1
delta_0 = 1
alpha = 136
beta = 0.02

N_samps = 10000
alpha_curr = rnorm(n,a_0,b_0)
beta_curr = rnorm(c_0,d_0)
nu_curr = NA

#making containers for output
alpha_out = beta_out = nu_out = rep(NA,N_samps)

#Looping for the sampler
for(i in 1:N_samps){
  #updating
  #the a,b,c,d,delta,zeta are all defined in terms of the statistics and the current values
  #in their definitions in q1e, where you see and alpha, beta, or nu 
  #you should use alpha_curr, beta_curr, or nu_curr
  
  #changing order to update nu first because you only were given alpha and beta starting values
  delta_curr = n/2 + delta_0
  zeta_curr = ((n*alpha^2 + w_n*beta^2 + u_n - 2*alpha*s_n - 2*beta*r_n + 2*alpha*beta*t_n)/2)+zeta_0
  nu_curr = rinvgamma(1, delta_curr, zeta_curr)

  b_curr = 1/(n/nu_curr+1/b_0)
  a_curr = ((s_n-beta*t_n)/nu_curr)/(n/(nu_curr+1)/b_0)
  alpha_curr = rnorm(1,a_curr,sqrt(b_curr)) 
  
  c_curr = (r_n - alpha*t_n)/nu_curr+c_0/d_0 / w_n/nu_curr+1/d_0
  d_curr = 1/w_n/nu_curr+1/d_0
  beta_curr = rnorm(1,c_curr,sqrt(d_curr)) 
  
      #storing
  nu_out[i] = nu_curr
  alpha_out[i] = alpha_curr
  beta_out[i] = beta_curr
}

plot(nu_out,pch='.')
acf(nu_out)
plot(alpha_out,pch='.')
acf(alpha_out)
plot(beta_out,pch='.')
acf(beta_out)

#Density plots
plot(density(nu_out))
plot(density(alpha_out))
plot(density(beta_out))
#posterior means
draws_nu_mean = mean(nu_out)
draws_alpha_mean = mean(alpha_out)
draws_beta_mean = mean(beta_out)
#intervals
quantile(nu_out,c(0.025,0.975))
quantile(alpha_out,c(0.025,0.975))
quantile(beta_out,c(0.025,0.975))
#correlation of parameters
cor(nu_out,alpha_out)
cor(nu_out,beta_out)
cor(alpha_out,beta_out)

#predictions
y_pred = rnorm(N_samps,draws_alpha_mean+x*draws_beta_mean,sqrt(draws_nu_mean))
#comparison
plot(density(y))
lines(density(y_pred),col="blue",lty=2)
legend("topright",legend=c("KDE-Data","KDE-Predictions"),col=c("black","blue"),lty=c(1,2))
```
All of the trace plots are kind of bad. Our acf's aren't great either. Maybe I've done something incorrect, or this is our first sign that the model is not good. 
I am really doing something here, feels very wrong. 

```{r,eval=FALSE}
y_pred = rnorm(n,draws_alpha_mean+x*draws_beta_mean,sqrt(draws_nu_mean))
plot(y_pred~x)
points(y~x)

x_new = 3800
y_pred = rnorm(N_samps,alpha_out+x_new*beta_out,sqrt(nu_out)) 
plot(density(y_pred))
lines(density(y[x>=3700 & x<=3900]),col="blue",lty=2)
```
Really not sure if this is doing a good job, I may have gotten a little lost in the weeds on this problem

