---
title: "hw7"
author: "Morgan Fissel"
date: "3/11/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Question 1
```{r}
analyze_two_by_two_table = function(x,y=NULL,signif_level=0.01){
  ##################
  ## For Task 3.0 ##
  ##################
  # if x is not a matrix, it treats x and y as factor variables
  
  if("matrix" %in% class(x)){
    N = x
  }else if("table" %in% class(x)){
    N = matrix(x,2,2)
    dimnames(N) = dimnames(x)
  }else{
    N = table(x,y)
  }
  
  ##################
  ## End Task 3.0 ##
  ##################
  
  #################################
  ## Start Functions for fitting ##
  #################################
  # now we input the MLE functions
  # this is just copying and pasting from above
  mle_two_by_two_table_full = function(N){
    # N is a 2x2 table of counts
    # log_odds_ratio are for log(p_{11}p_{22}/(p_{12}p_{21}))
    joint_probs = N/sum(N)
    row_probs = rowSums(joint_probs)
    col_probs = colSums(joint_probs)
    log_odds_ratio = log(joint_probs[1,1])-log(joint_probs[1,2])+
      log(joint_probs[2,2])-log(joint_probs[2,1])
    theta = list(joint_probs = joint_probs,
               row_probs = row_probs,
               col_probs = col_probs,
               log_odds_ratio = log_odds_ratio)
    loglik = sum(N*log(joint_probs))
    return(list(theta=theta,loglik=loglik))
  }
  mle_two_by_two_table_independent = function(N){
    # N is a 2x2 table of counts
    # log_odds_ratio are for log(p_{11}p_{22}/(p_{12}p_{21}))
    row_probs = rowSums(N)/sum(N)
    col_probs = colSums(N)/sum(N)
    joint_probs = outer(row_probs,col_probs,"*")
    log_odds_ratio = 0
    theta = list(joint_probs = joint_probs,
               row_probs = row_probs,
               col_probs = col_probs,
               log_odds_ratio = log_odds_ratio)
    loglik = sum(N*log(joint_probs))
    return(list(theta=theta,loglik=loglik))
  }
  mle_two_by_two_table_fixed_log_odds_ratio = function(N,log_odds_ratio_0){
    # N is a 2x2 table of counts
    # log_odds_ratio are for log(p_{11}p_{22}/(p_{12}p_{21}))
    f = function(theta_trans){
      col_probs = 1/(1+exp(c(-1,1)*theta_trans[1]))
      joint_probs = matrix(NA,2,2)
      col_1_cond_probs = 1/(1+exp(c(-1,1)*theta_trans[2]))
      joint_probs[,1] = col_probs[1]*col_1_cond_probs
      col_2_cond_probs = c(NA,NA)
      col_2_cond_probs[1] = col_1_cond_probs[1]/
        (col_1_cond_probs[2]*exp(log_odds_ratio_0)+col_1_cond_probs[1])
      col_2_cond_probs[2] = 1-col_2_cond_probs[1]
      joint_probs[,2] = col_probs[2]*col_2_cond_probs
      joint_probs[,2] = col_probs[2]*col_2_cond_probs
      loglik = sum(N*log(joint_probs))
    }
    control = list(fnscale=-1,maxit=1000,reltol=1e-12)
    opt = optim(c(0,0),f,control=control)
    opt_trans = opt$par
    col_probs = 1/(1+exp(c(-1,1)*opt_trans[1]))
    joint_probs = matrix(NA,2,2)
    col_1_cond_probs = 1/(1+exp(c(-1,1)*opt_trans[2]))
    joint_probs[,1] = col_probs[1]*col_1_cond_probs
    col_2_cond_probs = c(NA,NA)
    col_2_cond_probs[1] = col_1_cond_probs[1]/
      (col_1_cond_probs[2]*exp(log_odds_ratio_0)+col_1_cond_probs[1])
    col_2_cond_probs[2] = 1-col_2_cond_probs[1]
    joint_probs[,2] = col_probs[2]*col_2_cond_probs
    row_probs = rowSums(joint_probs)
    col_probs = colSums(joint_probs)
    log_odds_ratio = log_odds_ratio_0
    theta = list(joint_probs = joint_probs,
               row_probs = row_probs,
               col_probs = col_probs,
               log_odds_ratio = log_odds_ratio)
    loglik = opt$value
    return(list(theta=theta,loglik=loglik))
  }
  ###############################
  ## End Functions for fitting ##
  ###############################
  
  
  # now we start doing new things
  # for now, we are going to use meaningless placeholders
  # you will replace these placeholders later
  
  ##################
  ## For Task 3.1 ##
  ##################
  
  # replace with actual fit under HA
  fit_HA = mle_two_by_two_table_full(N)
  # replace with actual fit under H0
  fit_H0 = mle_two_by_two_table_independent(N)
  # replace with actual LRT stat
  lambda = 2*(fit_HA$loglik - fit_H0$loglik)
  cutoff = qchisq(1 - signif_level,1)
  decision = ifelse(lambda>cutoff,"Reject H_0: Independence","Fail to reject H_0: Independence")
  AIC_H0_minus_HA = lambda-2
  p_val = pchisq(lambda,1,lower.tail = FALSE)
  ##################
  ## End Task 3.1 ##
  ##################
  
  ##################
  ## For Task 3.2 ##
  ##################
  # confidence interval
  lambda_fun = function(log_odds_ratio_0=0,N,signif_level=0.01){
    fit_H0 = mle_two_by_two_table_fixed_log_odds_ratio(N,log_odds_ratio_0)
    fit_HA = mle_two_by_two_table_full(N)
    lambda = 2*(fit_HA$loglik - fit_H0$loglik) #lrt
    cutoff = qchisq(1 - signif_level,1)
    decision = ifelse(lambda>cutoff,"Reject H_0: Independence","Fail to reject H_0: Independence")
    AIC_H0_minus_HA = lambda-2
    p_val = pchisq(lambda,1,lower.tail = FALSE)
    if(p_val <= 0){
      return(Inf)
    }else{
      return(lambda-cutoff)
    }
  }


  ##################
  ## End Task 3.2 ##
  ##################
  
  ##################
  ## For Task 3.3 ##
  ##################
  # confidence interval
  lambda_fun = function(log_odds_ratio_0=0,N=x,signif_level=0.01){
    fit_H0 = mle_two_by_two_table_fixed_log_odds_ratio(N,log_odds_ratio_0)
    fit_HA = mle_two_by_two_table_full(N)
    lambda = 2*(fit_HA$loglik - fit_H0$loglik) #lrt
    cutoff = qchisq(1 - signif_level,1)
    decision = ifelse(lambda>cutoff,"Reject H_0: Independence","Fail to reject H_0: Independence")
    AIC_H0_minus_HA = lambda-2
    p_val = pchisq(lambda,1,lower.tail = FALSE)
    if(p_val <= 0){
      return(Inf)
    }else{
      lrt_stat_minus_cutoff = lambda-cutoff
      return(lrt_stat_minus_cutoff)
    }
  }

  theta_hat = mle_two_by_two_table_full(N)$theta$log_odds_ratio
  interval = c(theta_hat-1,theta_hat) #for the lower bound, using the constraint and the MLE
  lb = uniroot(lambda_fun,interval,extendInt = "downX",tol = 1e-10,maxiter=10000) #extendInt="downX" is       actually doing nothing here becasue of the constraint
  interval = c(theta_hat,theta_hat+1) #for the upper bound, MLE to MLE+10 (+10 was arbitrary and I could probably have been more clever    in how I chose it)
  ub = uniroot(lambda_fun,interval,extendInt = "upX",tol = 1e-10,maxiter=10000) #extendint="upX" is           extending the interval looking for an up crossing. This is in case the +10 was not enough.
  #return(c(estimate=theta_hat,lb=lb$root,ub=ub$root))

  

  ##################
  ## End Task 3.3 ##
  ##################
  
  return(list(
    fit_HA=fit_HA,
    fit_H0=fit_H0,
    lambda=lambda,
    AIC_H0_minus_HA=AIC_H0_minus_HA,
    cutoff=cutoff,
    decision=decision,
    p_val=p_val,
    signif_level=signif_level,
    conf_level=1-signif_level,
    conf_int=c(lb,ub)
    ))
  }
```
```{r}
UCB_agg_over_dept = apply(UCBAdmissions,c(1,2),sum)
analyze_two_by_two_table(UCB_agg_over_dept)
```
a)
p-val = 4.167175e-22
extremely small/close to 0.
b)
conf_int$root 
[1] 0.7756045

```{r}
dept_A = UCBAdmissions[,,'A']
analyze_two_by_two_table(dept_A)

dept_B = UCBAdmissions[,,'B']
analyze_two_by_two_table(dept_B)

dept_C = UCBAdmissions[,,'C']
analyze_two_by_two_table(dept_C)

dept_D = UCBAdmissions[,,'D']
analyze_two_by_two_table(dept_D)

dept_E = UCBAdmissions[,,'E']
analyze_two_by_two_table(dept_E)

```
c)
p_vals = (1.270705e-05, A)
          (0.611054, B)
          (0.3861648, C)
          (0.585223, D)
          (0.319648, E)
        
d)
Intervals above
e)
There is quite a difference in agg p_val, vs. non-agg, seems like there is inconsistency
between the aggregate data, and the non-aggregate data. Not sure what is causing this, but I would assume it has something to do with the sum, we use for the agg.

Question 2
```{r}
dept_A = UCBAdmissions[,,'A']
analyze_two_by_two_table(dept_A,signif_level = 0.01/6)

dept_B = UCBAdmissions[,,'B']
analyze_two_by_two_table(dept_B,signif_level = 0.01/6)

dept_C = UCBAdmissions[,,'C']
analyze_two_by_two_table(dept_C,signif_level = 0.01/6)

dept_D = UCBAdmissions[,,'D']
analyze_two_by_two_table(dept_D,signif_level = 0.01/6)

dept_E = UCBAdmissions[,,'E']
analyze_two_by_two_table(dept_E,signif_level = 0.01/6)
```
b) conclusions may change, but not by much.
```{r}
dept_A = UCBAdmissions[,,'A']
analyze_two_by_two_table(dept_A,signif_level = (1-0.99)^(1/6))

dept_B = UCBAdmissions[,,'B']
analyze_two_by_two_table(dept_B,signif_level = (1-0.99)^(1/6))

dept_C = UCBAdmissions[,,'C']
analyze_two_by_two_table(dept_C,signif_level = (1-0.99)^(1/6))

dept_D = UCBAdmissions[,,'D']
analyze_two_by_two_table(dept_D,signif_level = (1-0.99)^(1/6))

dept_E = UCBAdmissions[,,'E']
analyze_two_by_two_table(dept_E,signif_level = (1-0.99)^(1/6))
```
c) Once again everything stays pretty much the same.
